---
title: 'Homework 4: SCC and Climate Policy'
author: "Iris Foxfoot"
date: "5/19/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
options(scipen=999)
library(here)
library(tidyverse)
library(patchwork)
```

1. Using damages.csv, estimate a quadratic damage function relating the dollar value of damages to the change in global mean temperature. Omit an intercept term; damages by construction must equal zero when there is no climate change. Plot your estimated damage function, overlaid with a scatterplot of the underlying data.

$$ damages = -3018856250104(warming) + 19589020155100(warming)^2$$
```{r}
#read in data
damages_data <- read_csv(here("damages.csv"))

warming_data <- read_csv(here("warming.csv"))

#creating vectors of values
damages_data <- damages_data %>% 
  mutate(warming2 = warming^2)

#creating vectors
damages <- damages_data$damages
warming <- damages_data$warming
warming2 <- damages_data$warming2

#estimating quadratic model
damages_quad_model=lm(method = 'lm', damages ~ warming + warming2)

summary(damages_quad_model)

#creating a sequence of warming values to test predictions on
warmingvalues <- seq(0.6900000, 9.243000, 0.1)

#generating predictions
predictedcounts <- predict(damages_quad_model,list(warming=warmingvalues, warming2=warmingvalues^2))

#plotting
plot(warming, damages, pch=16, xlab = "Warming (Celsius)", ylab = "damage in $", cex.lab = 1.3, col = "blue");
lines(warmingvalues, predictedcounts, col = "red", lwd = 3)
```

2. Use warming.csv and your estimated damage function to predict damages in each year under the baseline climate and the pulse scenario. Make four plots: (1) damages over time without the pulse, (2) damages over time with the pulse, (3) the difference in damages over time that arises from the pulse, and (4) the difference in damages over time from the pulse per ton of
CO2 (you can assume that each ton of the pulse causes the same amount of damage).

```{r}
#making a damage function--do it with stored intercepts later
damage_function <- function(x){
  damage <- -3018856250104*x + 19589020155100*x^2
  return(damage)
}

#creating new columns
warming_scenarios <- warming_data %>% 
  mutate(damage_no_pulse = damage_function(warming_baseline)) %>% 
  mutate(damage_with_pulse = damage_function(warming_pulse)) %>% 
  mutate(pulse_damages = damage_with_pulse - damage_no_pulse) %>% 
  mutate(pulse_damages_per_ton = pulse_damages/35000000000)

#create graphs
damage_no_pulse_graph <- ggplot(data = warming_scenarios, aes(x = year, y = damage_no_pulse)) +
  geom_point()

damage_with_pulse_graph <- ggplot(data = warming_scenarios, aes(x = year, y = damage_with_pulse)) +
  geom_point()

pulse_damages_graph <- ggplot(data = warming_scenarios, aes(x = year, y = pulse_damages)) +
  geom_point()

pulse_damages_per_ton_graph <- ggplot(data = warming_scenarios, aes(x = year, y = pulse_damages_per_ton)) +
  geom_point()
 
#put the graphs together
(damage_no_pulse_graph + damage_with_pulse_graph)/(pulse_damages_graph + pulse_damages_per_ton_graph)
```

3. The SCC is the present discounted value of the stream of future damages caused by one additional ton of CO2. The Obama Administration used a discount rate of 3% to discount damages. Recently, New York State used a discount rate of 2%. Calculate and make a plot of the SCC (y-axis) against the discount rate (x-axis) for a reasonable range of discount rates.

$ SCC = \sum_ 1/(1 +r)^t $ Fix formula...haha

```{r}
#function that takes in two vectors
#num = first co2 damage per ton
#dr = discount rate
scc_function <- function(num, dr = .02) {
  sum <- 0;
  r <- dr
  for(t in c(1:length(num))){
      present <- num[t]/(1+r)^t
      sum <- sum+present
  }
  return(sum)
}

#generating a list of discount rates to use
discount_rate <- seq(.01, .1, .001)

#testing it out
scc_function(warming_scenarios$pulse_damages_per_ton, discount_rate)

#creating a dataframe
scc <- data.frame(discount_rate<-discount_rate, 
scc<-scc_function(warming_scenarios$pulse_damages_per_ton, discount_rate))
colnames(scc) = c("discount_rate", "scc")

#graphing
ggplot(data = scc, aes(x = discount_rate, y = scc)) +
  geom_point()
```

4. The National Academies of Sciences, Engineering, and Medicine advised the government in a 2017 report to use the Ramsey Rule when discounting within the SCC calculation:

r = ρ + ηg

Using ρ = 0.001, η = 2, and g = 0.01, what is the SCC? Locate this point on your graph from above.

```{r}
#finding discount rate
ramsey_discount_rate <- 0.001 + 2*0.01

#inserting this into discount rate
ramsey_point <- data.frame(ramsey_discount_rate, scc_function(warming_scenarios$pulse_damages_per_ton, ramsey_discount_rate))

#changing col names
colnames(ramsey_point) = c("discount_rate", "ramsey_scc")

ggplot() +
  geom_point(data = scc, aes(x = discount_rate, y = scc)) +
  geom_point(data = ramsey_point, aes(x = ramsey_discount_rate, y = ramsey_scc, color = "red", size = 1))#
```

5. Now suppose there are two possible climate policies that can be pursued. Policy A is business as usual and Policy B is to take immediate and strong action on climate change. Use these facts

• If you undertake Policy A there are two possible outcomes. Either warming will occur as in the “baseline” (i.e. “no-pulse”) dataset above (this happens with probability 0.5) or warming each year will be 1.5 times that in the “baseline” dataset (with probability 0.5).

• Under Policy B, warming will continue until 2050 as in the “baseline” dataset, and then will stabilize at 1.29 degrees and stay that way forever.

• Society is risk neutral

• Use a discount rate of 2% What is the expected present value of damages up to 2100 under Policy A? What is the expected present value of damages up to 2100 under Policy B? Suppose undertaking Policy A costs zero and undertaking Policy B costs X. How large could X be for it to still make economic sense to pursue Policy B instead of Policy A? Qualitatively, how would your answer change if society were risk averse?


